---
title: "RFMO Data Cleaning - IOTC"
output: html_notebook
---

This is a notebook designed to grab data saved in Google Drive and re-format it for easier analyses later.

# Table of Contents: {#top}

- [CE Coastal]
- [CE Longline]
- [CE Surface]
- [NCDB]
- [SFSKH]
- [Combining Files]

```{r, include=FALSE}
# Import libraries
library(tidyverse)
library(reshape)
library(readxl)
```

# Housekeeping
```{r, echo=TRUE}
# List files to choose from
file_path = "../../data/rfmo-observer-data/inputs/iotc/"
save_location = "../../data/rfmo-observer-data/intermediates/iotc/"
outputs_location = "../../data/rfmo-observer-data/outputs/"
```

# Load Data
```{r, echo=TRUE}
# List csv files
list_files <- list.files(file_path, pattern="csv")

# Load files
for(file in list_files) { 
  name <- colsplit(file, "[.]", c("",""))[,1]
  name <- as.character(colsplit(name, "-", c("","","",""))[,4])
  temp <- read.csv(paste0(file_path, file))
  assign(name, temp)
}

# Clean up
remove(temp)

# Load excel files
ncdb <- read_xlsx(paste0(file_path, "IOTC-2019-DATASETS-NCDB_061219.xlsx"), sheet="Catches_Captures")
sfskh <- read_xlsx(paste0(file_path, "IOTC-2019-WPEB15-DATA09 - SFSKH.xlsx"), sheet="SKH")

# Fleet/Country
fleet <- read_xlsx(paste0(file_path, "IOTC-2019-DATASETS-CEref.xlsx"), sheet="CodeFleet")
colnames(fleet) <- c("Fleet", "Country", "French")

# Gear
gear <- read_xlsx(paste0(file_path, "IOTC-2019-DATASETS-CEref.xlsx"), sheet="CodeGear")
colnames(gear)[4] <- "gear"

# Effort
effort <- read_xlsx(paste0(file_path, "IOTC-2019-DATASETS-CEref.xlsx"), sheet="CodeEffort")
effort <- unique(effort[,c("EffortUnits", "Description")])

# Grid sizes
grid_size <- data.frame("size" = c(1,2,3,4,5,6), "lat_shift" = c(5,10,10,20,1,5), "lon_shift"=c(10,20,10,20,1,5))
grid_size <- grid_size %>% # add spatial notes
  mutate(spatial_notes = paste0("center of ", lat_shift, "x", lon_shift, " cell"))
grid_size <- grid_size %>%  # fix shift (right now shift == grid size, when it should be 0.5*grid size)
  mutate(lat_shift = lat_shift/2, 
         lon_shift = lon_shift/2)

# Species
species <- read.delim(paste0(file_path, "ASFIS_sp_2019.txt"), sep=",")

# NCDB Location Info
ncdb_gps <- data.frame("AreaIOTC" = c("Western Indian Ocean", "Eastern Indian Ocean"), 
                       "latitude" = c(-25, -30), 
                       "longitude" = c(60, 90), 
                       "spatial_notes" = c("generated point within larger area",
                                           "generated point within larger area"))
```
[Back To Top](#top)

# CE Coastal
```{r, echo=TRUE}
# Filter to year 2000 or greater
CECoastal <- CECoastal %>% 
  filter(Year >= 2000)

# Pivot tables so each row represents a different species
CECoastal <- CECoastal %>% 
  pivot_longer(cols=YFT.NO:NTAD.MT, names_to = 'speciescode', values_to = 'catch') %>% 
  filter(catch > 0)

# Split speciescodes into species code and catch units (number vs metric tonnes)
CECoastal <- CECoastal %>% 
  separate(speciescode, into=c("speciescode", "catch_units"), sep=-3) %>% 
  mutate(catch_units = gsub(".NO", "count", catch_units), 
         catch_units = gsub(".MT", "metric tonnes", catch_units))

# Fix GPS locations and add spatial notes
CECoastal <- CECoastal %>% 
  separate(Grid, into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Add fleet/country
CECoastal <- CECoastal %>% 
  mutate(Fleet = gsub(" ", "", Fleet)) %>% # fix extra spaces
  left_join(fleet[,c("Fleet", "Country")], by="Fleet") %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country))

# Add species data
CECoastal <- left_join(CECoastal, species[,c("X3A_CODE", "Scientific_name", "English_name")], by = c("speciescode" = "X3A_CODE") )

# Make all species names all caps
CECoastal$species_sciname <- toupper(CECoastal$Scientific_name)
CECoastal$species_commonname <- toupper(CECoastal$English_name)

# Fix species names that were not included in the ASFIS codes
CECoastal <- CECoastal %>% 
  mutate(species_commonname = ifelse(speciescode == "BILL", "BILLFISH NEI",
                                     ifelse(speciescode == "NTAD", "OTHER SPECIES", species_commonname)), 
         species_sciname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_sciname))

# Add gear
CECoastal <- CECoastal %>% 
  left_join(gear[,c("CECode", "gear")], by=c("Gear" = "CECode")) %>% 
  mutate(gear = tolower(gear),
         gear = gsub("  ", " ", gear))

# Add effort units
CECoastal <- CECoastal %>% 
  left_join(effort[,c("EffortUnits", "Description")], by="EffortUnits") %>%
  mutate(Description = gsub("Number of ", "", Description), 
         Description = gsub("Men-days", "men days", Description))

# Create/mutate relevant columns
CECoastal <- CECoastal %>% 
  mutate(rfmo = "IOTC", 
         year = Year, 
         month = MonthStart, 
         time_period = "monthly", 
         country = Country,
         settype = NA,  
         effort = Effort,
         effort_units = Description,
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA,
         is_targeted = ifelse(grepl("TUNA", species_commonname), "yes-implied", "no"),
         quality_code = QualityCode, 
         was_generated = "no", 
         sources = NA)

# Select relevant columns 
CECoastal <- CECoastal[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname", "is_targeted", "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- CECoastal %>% 
  filter(time_period == "monthly") %>%
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CECoastal), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Generate yearly data (aggregates)
yearly_agg <- CECoastal %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CECoastal), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Combine raw with generated data
CECoastal <- CECoastal %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "CECoastal.csv"
write.csv(CECoastal, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(CECoastal, quart_agg, yearly_agg)
```
[Back To Top](#top)

# CE Longline
```{r, echo=TRUE}
# Filter to year 2000 or greater
CELongline <- CELongline %>% 
  filter(Year >= 2000)

# Pivot tables so each row represents a different species
CELongline <- CELongline %>% 
  pivot_longer(cols=YFT.NO:NTAD.MT, names_to = 'speciescode', values_to = 'catch') %>% 
  filter(catch > 0)

# Split speciescodes into species code and catch units (number vs metric tonnes)
CELongline <- CELongline %>% 
  separate(speciescode, into=c("speciescode", "catch_units"), sep=-3) %>% 
  mutate(catch_units = gsub(".NO", "count", catch_units), 
         catch_units = gsub(".MT", "metric tonnes", catch_units))

# Fix GPS locations and add spatial notes
CELongline <- CELongline %>% 
  separate(Grid, into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Add fleet/country
CELongline <- CELongline %>% 
  mutate(Fleet = gsub(" ", "", Fleet)) %>% # fix extra spaces 
  left_join(fleet[,c("Fleet", "Country")], by="Fleet") %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country), 
         Country = gsub(",China", "", Country))

# Add species data
CELongline <- left_join(CELongline, species[,c("X3A_CODE", "Scientific_name", "English_name")], by = c("speciescode" = "X3A_CODE") )

# Make all species names all caps
CELongline$species_sciname <- toupper(CELongline$Scientific_name)
CELongline$species_commonname <- toupper(CELongline$English_name)

# Fix species names that were not included in the ASFIS codes
CELongline <- CELongline %>% 
  mutate(species_commonname = ifelse(speciescode == "BILL", "BILLFISH NEI",
                                     ifelse(speciescode == "NTAD", "OTHER SPECIES", species_commonname)), 
         species_sciname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_sciname))

# Add gear
CELongline <- left_join(CELongline, gear[,c("CECode", "gear")], by=c("Gear" = "CECode")) %>%
  mutate(gear = tolower(gear), 
         gear = gsub("  ", " ", gear))

# Add effort units
CELongline <- CELongline %>% 
  left_join(effort[,c("EffortUnits", "Description")], by="EffortUnits") %>%
  mutate(Description = gsub("Number of ", "", Description), 
         Description = gsub("Men-days", "men days", Description), 
         Description = gsub("sets", "number of sets", Description))

# Create/mutate relevant columns
CELongline <- CELongline %>% 
  mutate(rfmo = "IOTC", 
         year = Year, 
         month = MonthStart, 
         time_period = "monthly", 
         country = Country,
         settype = NA,  
         effort = Effort,
         effort_units = Description,
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA,
         is_targeted = ifelse(grepl("TUNA", species_commonname), "yes-implied", "no"),
         quality_code = QualityCode, 
         was_generated = "no", 
         sources = NA)

# Select relevant columns 
CELongline <- CELongline[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname", "is_targeted", "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- CELongline %>% 
  filter(time_period == "monthly") %>% 
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CELongline), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Generate yearly data (aggregates)
yearly_agg <- CELongline %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CELongline), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Combine raw with generated data
CELongline <- CELongline %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "CELongline.csv"
write.csv(CELongline, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(CELongline, quart_agg, yearly_agg)
```
[Back To Top](#top)

# CE Surface
```{r, echo=TRUE}
# Filter to year 2000 or greater
CESurface <- CESurface %>% 
  filter(Year >= 2000)

# Pivot tables so each row represents a different species
CESurface <- CESurface %>% 
  pivot_longer(cols=YFT.FS:NTAD.UNCL, names_to = 'speciescode', values_to = 'catch') %>% 
  filter(catch > 0)

# Split speciescodes into species code and settype
CESurface <- CESurface %>% 
  separate(speciescode, into=c("speciescode", "settype"), sep="[.]") %>% 
  mutate(settype = gsub("FS", "free schools", settype), 
         settype = gsub("LS", "associated schools", settype), 
         settype = gsub("UNCL", NA, settype))

# Fix GPS locations and add spatial notes
CESurface <- CESurface %>% 
  separate(Grid, into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Add fleet/country
CESurface <- CESurface %>% 
  mutate(Fleet = gsub(" ", "", Fleet)) %>% # fix extra spaces 
  left_join(fleet[,c("Fleet", "Country")], by="Fleet") %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country), 
         Country = gsub(",China", "", Country))

# Add species data
CESurface <- left_join(CESurface, species[,c("X3A_CODE", "Scientific_name", "English_name")], by = c("speciescode" = "X3A_CODE") )

# Make all species names all caps
CESurface$species_sciname <- toupper(CESurface$Scientific_name)
CESurface$species_commonname <- toupper(CESurface$English_name)

# Fix species names that were not included in the ASFIS codes
CESurface <- CESurface %>% 
  mutate(species_commonname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_commonname),
    species_sciname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_sciname))

# Add gear
CESurface <- left_join(CESurface, gear[,c("CECode", "gear")], by=c("Gear" = "CECode")) %>%
  mutate(gear = tolower(gear), 
         gear = gsub("  ", " ", gear), 
         gear = gsub("purse seine", "purse-seine", gear),
         gear = ifelse(Gear == "BBOF", "baitboat (offshore)", gear))

# Add effort units
CESurface <- CESurface %>% 
  left_join(effort[,c("EffortUnits", "Description")], by="EffortUnits") %>%
  mutate(Description = gsub("Number of ", "", Description), 
         Description = gsub("sets", "number of sets", Description))

# Create/mutate relevant columns
CESurface <- CESurface %>% 
  mutate(rfmo = "IOTC", 
         year = Year, 
         month = MonthStart, 
         time_period = "monthly", 
         country = Country,
         effort = Effort,
         effort_units = Description,
         catch_units = "metric tonnes",
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA, 
         is_targeted = ifelse(grepl("TUNA", species_commonname), "yes-implied", "no"),
         quality_code = QualityCode, 
         was_generated = "no", 
         sources = NA)

# Select relevant columns 
CESurface <- CESurface[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname", "is_targeted", "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- CESurface %>% 
  filter(time_period == "monthly") %>%
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CESurface), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Generate yearly data (aggregates)
yearly_agg <- CESurface %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CESurface), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Combine raw with generated data
CESurface <- CESurface %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "CESurface.csv"
write.csv(CESurface, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(CESurface, quart_agg, yearly_agg)
```
[Back To Top](#top)

# NCDB 
Not sure if this dataset is relevant; no effort data are provided. 
```{r, echo=TRUE}
# Keep only data from the 2000 onward
ncdb <- ncdb %>% 
  filter(`Year/An` >= 2000)

# Merge with area information to get gps locations  
ncdb <- ncdb %>% 
  left_join(ncdb_gps, by=c("AreaIOTC")) 

# Make species/common names consistent 
ncdb$species_commonname <- toupper(ncdb$Species)
ncdb$species_sciname <- toupper(ncdb$SpLat)

# Create/alter columns 
ncdb <- ncdb %>% 
  mutate(rfmo = "IOTC", 
         year = `Year/An`, 
         month = NA, 
         time_period = "yearly",
         country = str_to_title(Fleet),
         gear = tolower(Gear),
         settype = NA, 
         effort = NA, 
         effort_units = NA, 
         is_targeted = ifelse(SpIOTC == TRUE, "yes-stated", "no"),
         catch = `Catch/Capture(t)`,
         catch_units = "metric tonnes",
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA,
         quality_code = NA, 
         was_generated = "no", 
         sources = NA
         )

# Keep only relevant columns
ncdb <- ncdb[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname", "is_targeted", "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Fix gear types to work with other RFMOs
ncdb <- ncdb %>% 
  mutate(gear = gsub("purse seine", "purse-seine", gear))

# Fix countries
ncdb <- ncdb %>% 
  mutate(country = gsub("Eu.", "", country), 
         country = gsub("Un.", "United", country), 
         country = gsub("Rep.", "Republic", country), 
         country = gsub(",China", "", country),
         country = gsub("[.]", " ", country), 
         country = str_to_title(country), 
         country = gsub("Uk", "UK", country))

# Save File
filename <- "ncdb.csv"
write.csv(ncdb, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(ncdb)
```
[Back To Top](#top)

# SFSKH
It's unclear if these data are relevant. Provides size estimates for sharks caught, but there's no error estimates. 
Will add anyway, and include another column in the new dataset (will have to update other RFMOs).
```{r, echo=TRUE}
# Fix GPS locations and add spatial notes
sfskh <- sfskh %>% 
  separate(Grid, into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Move size classes to different columns
sfskh <- sfskh %>% 
  pivot_longer(cols=C001:C150, names_to = 'class_code', values_to = 'catch') %>% 
  filter(catch > 0)

# Fix class codes to reflect actual sizes
sfskh <- sfskh %>% 
  mutate(class_code = gsub("C", "", class_code),
         class_code = as.numeric(class_code),
         measurement = FirstClassLow+(SizeInterval*(class_code-1)), 
         measurement_method = "fork length", 
         measurement_unit = "cm")

# Get country codes
sfskh <- sfskh %>% 
  left_join(fleet[,c("Fleet", "Country")], by="Fleet") %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country), 
         Country = gsub(",China", "", Country))

# Get gear codes
sfskh <- sfskh %>% 
  left_join(gear[,c("CECode", "gear")], by=c("Gear" = "CECode")) %>%
  mutate(gear = tolower(gear), 
         gear = gsub("  ", " ", gear), 
         gear = ifelse(Gear == "LLOB", "longline", gear), # not in the metadata - guessing
         gear = ifelse(Gear == "ELLOB", "longline targeting swordfish", gear), 
         gear = ifelse(Gear == "PSOB", "purse-seine", gear))

# Get species
sfskh <- sfskh %>% 
  left_join(species[,c("X3A_CODE", "Scientific_name", "English_name")], by = c("Species" = "X3A_CODE") )

# Alter settype 
sfskh <- sfskh %>% 
  mutate(settype = SchoolType, 
         settype = gsub("UNCL", "unclassified", settype), 
         settype = gsub("FS", "free schools", settype), 
         settype = gsub("LS", "associated schools", settype))

# Add other columns
sfskh <- sfskh %>% 
  mutate(rfmo = "IOTC",
         year = Year, 
         month = MonthStart, 
         time_period = MonthEnd - MonthStart, 
         time_period = gsub(0, "monthly", time_period), 
         time_period = gsub(6, "7 months", time_period), 
         time_period = gsub(2, "quarterly", time_period), 
         country = Country, 
         effort = NA, 
         effort_units = NA, 
         species_commonname = toupper(English_name), 
         species_sciname = toupper(Scientific_name),
         is_targeted = "no", 
         catch_units = "count",
         quality_code = 0, 
         was_generated = "no", 
         sources = NA)

# Keep only relevant columns
sfskh <- sfskh[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname", "is_targeted", "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- sfskh %>% 
  filter(time_period == "monthly") %>% 
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(sfskh), c("catch"))) %>%
  summarise(catch = sum(catch))

# Generate yearly data (aggregates)
yearly_agg <- sfskh %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(sfskh), c("catch"))) %>%
  summarise(catch = sum(catch))

# Combine raw with generated data
sfskh <- sfskh %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "sfskh.csv"
write.csv(sfskh, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(sfskh, quart_agg, yearly_agg)
```
[Back To Top](#top)

# Combining Files
```{r echo=TRUE}
# Find list of cleaned data
files_list <- list.files(save_location, pattern="csv")

# Make sure it doesn't include the 'all' dataset
files_list <- files_list[!grepl("all", files_list)]

# Dataframe to hold combined data
combined_dat <- NULL

# Loop through files and save to master dataframe
for(file in files_list) {
  temp <- read.csv(paste0(save_location, file))
  combined_dat <- rbind(combined_dat, temp)
}

# Remove instances of NA for lat or lon (some issues with grid in sfskh)
combined_dat <- combined_dat %>% 
  filter(!is.na(latitude) & !is.na(longitude))

# Save final data file for this RFMO
filename <- "iotc-all.csv"
write.csv(combined_dat, paste0(outputs_location, filename), row.names = FALSE)
```
[Back To Top](#top)