---
title: "RFMO Data Cleaning - IOTC"
output: html_notebook
---

This is a notebook designed to grab data and re-format it for easier analyses later.
Note that some of the species scientific names are not present, we will add these in later analyses so that we can ensure RFMOs have consistent species names. 

# Table of Contents: {#top}

- [CE Other]
- [CE Longline]
- [CE Surface]
- [Interactions]
- [Combining Files]

```{r, include=FALSE}
# Import libraries
library(tidyverse)
library(reshape)
library(readxl)
library(here)
```

# Housekeeping
```{r, echo=TRUE}
# List files to choose from
file_path <- file.path(here::here(), "data-updated/rfmo-data/inputs/iotc/")
save_location <- file.path(here::here(), "data-updated/rfmo-data/intermediates/iotc/")
outputs_location <- file.path(here::here(), "data-updated/rfmo-data/outputs/")
```

# Load Data
```{r, echo=TRUE}
# List csv files
list_files <- list.files(file_path, pattern="csv")

# Load files
for(file in list_files) { 
  name <- str_split(file, "[.]")[[1]][1]
  name <- gsub("_1950", "", as.character(str_split(name, "-")[[1]][6]))
  temp <- read.csv(paste0(file_path, file))
  assign(name, temp)
}

# Clean up
remove(temp)

# Fleet/Country
fleet <- read_xlsx(paste0(file_path, "IOTC-DATASETS-2022-02-23-CEref_1950-2020.xlsx"), sheet="CodeFleet")
colnames(fleet) <- c("Fleet", "Country", "French")

# Gear
gear <- read_xlsx(paste0(file_path, "IOTC-DATASETS-2022-02-23-CEref_1950-2020.xlsx"), sheet="CodeGear")
colnames(gear)[4] <- "gear"

# Effort
effort <- read_xlsx(paste0(file_path, "IOTC-DATASETS-2022-02-23-CEref_1950-2020.xlsx"), sheet="CodeEffort")
effort <- unique(effort[,c("EffortUnits", "Description")])

# add Interactions datast
interactions <- read_xlsx(paste0(file_path, "IOTC-2020-WPEB16-DATA12_CE.xlsx"), 
                          sheet = "CE", skip = 1)
interactions_spp <- read_xlsx(paste0(file_path, "IOTC-2020-WPEB16-DATA12_CE.xlsx"), 
                          sheet = "REF_SPECIES")

# Grid sizes
grid_size <- data.frame("size" = c(1,2,3,4,5,6), "lat_shift" = c(5,10,10,20,1,5), "lon_shift"=c(10,20,10,20,1,5))
grid_size <- grid_size %>% # add spatial notes
  mutate(spatial_notes = paste0("center of ", lat_shift, "x", lon_shift, " cell"))
grid_size <- grid_size %>%  # fix shift (right now shift == grid size, when it should be 0.5*grid size)
  mutate(lat_shift = lat_shift/2, 
         lon_shift = lon_shift/2)

# Species
species <- read.delim(paste0(file_path, "ASFIS_sp_2019.txt"), sep=",")
```
[Back To Top](#top)

# CE Other
```{r, echo=TRUE}
# Filter to year 2012 or greater
CEOther <- CEOther %>% 
  filter(Year >= 2012)

# Pivot tables so each row represents a different species
CEOther <- CEOther %>% 
  pivot_longer(cols=YFT.NO:NTAD.MT, names_to = 'speciescode', values_to = 'catch') %>% 
  filter(catch > 0)

# Split speciescodes into species code and catch units (number vs metric tonnes)
CEOther <- CEOther %>% 
  separate(speciescode, into=c("speciescode", "catch_units"), sep=-3) %>% 
  mutate(catch_units = gsub(".NO", "count", catch_units), 
         catch_units = gsub(".MT", "metric tonnes", catch_units))

# Fix GPS locations and add spatial notes
CEOther <- CEOther %>% 
  separate(Grid, into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Add fleet/country
CEOther <- CEOther %>% 
  mutate(Fleet = gsub(" ", "", Fleet)) %>% # fix extra spaces
  left_join(fleet[,c("Fleet", "Country")], by="Fleet") %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country))

# Add species data
CEOther <- left_join(CEOther, species[,c("X3A_CODE", "Scientific_name", "English_name")], by = c("speciescode" = "X3A_CODE") )

# Make all species names all caps
CEOther$species_sciname <- toupper(CEOther$Scientific_name)
CEOther$species_commonname <- toupper(CEOther$English_name)

# Fix species names that were not included in the ASFIS codes
CEOther <- CEOther %>% 
  mutate(species_commonname = ifelse(speciescode == "BILL", "BILLFISH NEI",
                                     ifelse(speciescode == "NTAD", "OTHER SPECIES", species_commonname)), 
         species_sciname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_sciname))

# Add gear
CEOther <- CEOther %>% 
  left_join(gear[,c("CECode", "gear")], by=c("Gear" = "CECode")) %>% 
  mutate(gear = tolower(gear),
         gear = gsub("  ", " ", gear))

# Add effort units
CEOther <- CEOther %>% 
  left_join(effort[,c("EffortUnits", "Description")], by="EffortUnits") %>%
  mutate(Description = gsub("Number of ", "", Description), 
         Description = gsub("Men-days", "men days", Description))

# Create/mutate relevant columns
CEOther <- CEOther %>% 
  mutate(rfmo = "IOTC", 
         year = Year, 
         month = MonthStart, 
         time_period = "monthly", 
         country = Country,
         settype = NA,  
         effort = Effort,
         effort_units = Description,
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA,
         quality_code = QualityCode, 
         was_generated = "no", 
         sources = "IOTC-DATASETS-2022-02-23-CEOther_1950-2020")

# Select relevant columns 
CEOther <- CEOther[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname",  "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- CEOther %>% 
  filter(time_period == "monthly") %>%
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CEOther), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Generate yearly data (aggregates)
yearly_agg <- CEOther %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CEOther), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Combine raw with generated data
CEOther <- CEOther %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "CEOther.csv"
write.csv(CEOther, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(CEOther, quart_agg, yearly_agg)
```
[Back To Top](#top)

# CE Longline
```{r, echo=TRUE}
# Filter to year 2012 or greater
CELongline <- CELongline %>% 
  filter(Year >= 2012)

# Pivot tables so each row represents a different species
CELongline <- CELongline %>% 
  pivot_longer(cols=YFT.NO:NTAD.MT, names_to = 'speciescode', values_to = 'catch') %>% 
  filter(catch > 0)

# Split speciescodes into species code and catch units (number vs metric tonnes)
CELongline <- CELongline %>% 
  separate(speciescode, into=c("speciescode", "catch_units"), sep=-3) %>% 
  mutate(catch_units = gsub(".NO", "count", catch_units), 
         catch_units = gsub(".MT", "metric tonnes", catch_units))

# Fix GPS locations and add spatial notes
CELongline <- CELongline %>% 
  separate(Grid, into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Add fleet/country
CELongline <- CELongline %>% 
  mutate(Fleet = gsub(" ", "", Fleet)) %>% # fix extra spaces 
  left_join(fleet[,c("Fleet", "Country")], by="Fleet") %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country), 
         Country = gsub(",China", "", Country))

# Add species data
CELongline <- left_join(CELongline, species[,c("X3A_CODE", "Scientific_name", "English_name")], by = c("speciescode" = "X3A_CODE") )

# Make all species names all caps
CELongline$species_sciname <- toupper(CELongline$Scientific_name)
CELongline$species_commonname <- toupper(CELongline$English_name)

# Fix species names that were not included in the ASFIS codes
CELongline <- CELongline %>% 
  mutate(species_commonname = ifelse(speciescode == "BILL", "BILLFISH NEI",
                                     ifelse(speciescode == "NTAD", "OTHER SPECIES", species_commonname)), 
         species_sciname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_sciname))

# Add gear
CELongline <- left_join(CELongline, gear[,c("CECode", "gear")], by=c("Gear" = "CECode")) %>%
  mutate(gear = tolower(gear), 
         gear = gsub("  ", " ", gear))

# Add effort units
CELongline <- CELongline %>% 
  left_join(effort[,c("EffortUnits", "Description")], by="EffortUnits") %>%
  mutate(Description = gsub("Number of ", "", Description), 
         Description = gsub("Men-days", "men days", Description), 
         Description = gsub("sets", "number of sets", Description))

# Create/mutate relevant columns
CELongline <- CELongline %>% 
  mutate(rfmo = "IOTC", 
         year = Year, 
         month = MonthStart, 
         time_period = "monthly", 
         country = Country,
         settype = NA,  
         effort = Effort,
         effort_units = Description,
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA,
         quality_code = QualityCode, 
         was_generated = "no", 
         sources = "IOTC-DATASETS-2022-02-23-CELongline_1950-2020")

# Select relevant columns 
CELongline <- CELongline[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname",  "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- CELongline %>% 
  filter(time_period == "monthly") %>% 
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CELongline), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Generate yearly data (aggregates)
yearly_agg <- CELongline %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CELongline), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Combine raw with generated data
CELongline <- CELongline %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "CELongline.csv"
write.csv(CELongline, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(CELongline, quart_agg, yearly_agg)
```
[Back To Top](#top)

# CE Surface
```{r, echo=TRUE}
# Filter to year 2012 or greater
CESurface <- CESurface %>% 
  filter(Year >= 2012)

# Pivot tables so each row represents a different species
CESurface <- CESurface %>% 
  pivot_longer(cols=YFT.FS:NTAD.UNCL, names_to = 'speciescode', values_to = 'catch') %>% 
  filter(catch > 0)

# Split speciescodes into species code and settype
CESurface <- CESurface %>% 
  separate(speciescode, into=c("speciescode", "settype"), sep="[.]") %>% 
  mutate(settype = gsub("FS", "free schools", settype), 
         settype = gsub("LS", "associated schools", settype), 
         settype = gsub("UNCL", NA, settype))

# Fix GPS locations and add spatial notes
CESurface <- CESurface %>% 
  separate(Grid, into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Add fleet/country
CESurface <- CESurface %>% 
  mutate(Fleet = gsub(" ", "", Fleet)) %>% # fix extra spaces 
  left_join(fleet[,c("Fleet", "Country")], by="Fleet") %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country), 
         Country = gsub(",China", "", Country))

# Add species data
CESurface <- left_join(CESurface, species[,c("X3A_CODE", "Scientific_name", "English_name")], by = c("speciescode" = "X3A_CODE") )

# Make all species names all caps
CESurface$species_sciname <- toupper(CESurface$Scientific_name)
CESurface$species_commonname <- toupper(CESurface$English_name)

# Fix species names that were not included in the ASFIS codes
CESurface <- CESurface %>% 
  mutate(species_commonname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_commonname),
    species_sciname = ifelse(speciescode == "NTAD", "OTHER SPECIES", species_sciname))

# Add gear
CESurface <- left_join(CESurface, gear[,c("CECode", "gear")], by=c("Gear" = "CECode")) %>%
  mutate(gear = tolower(gear), 
         gear = gsub("  ", " ", gear), 
         gear = gsub("purse seine", "purse-seine", gear),
         gear = ifelse(Gear == "BBOF", "baitboat (offshore)", gear))

# Add effort units
CESurface <- CESurface %>% 
  left_join(effort[,c("EffortUnits", "Description")], by="EffortUnits") %>%
  mutate(Description = gsub("Number of ", "", Description), 
         Description = gsub("sets", "number of sets", Description))

# Create/mutate relevant columns
CESurface <- CESurface %>% 
  mutate(rfmo = "IOTC", 
         year = Year, 
         month = MonthStart, 
         time_period = "monthly", 
         country = Country,
         effort = Effort,
         effort_units = Description,
         catch_units = "metric tonnes",
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA, 
         quality_code = QualityCode, 
         was_generated = "no", 
         sources = "IOTC-DATASETS-2022-02-23-CESurface_1950-2020")

# Select relevant columns 
CESurface <- CESurface[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname", "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- CESurface %>% 
  filter(time_period == "monthly") %>%
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CESurface), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Generate yearly data (aggregates)
yearly_agg <- CESurface %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(CESurface), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Combine raw with generated data
CESurface <- CESurface %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "CESurface.csv"
write.csv(CESurface, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(CESurface, quart_agg, yearly_agg)
```
[Back To Top](#top)

# Interactions
```{r}
# Filter to year 2012 or greater
interactions <- interactions %>% 
  filter(YEAR >= 2012)

# Fix GPS locations and add spatial notes
interactions <- interactions %>% 
  separate(GRID,  into = c("size", "quadrant", "latitude", "longitude"), sep = c(1,2,4), convert=TRUE) %>%
  left_join(grid_size, by="size") %>%
  mutate(latitude = ifelse(quadrant == 2|quadrant == 3, as.numeric(latitude)*(-1)-lat_shift, as.numeric(latitude)+lat_shift), 
         longitude == ifelse(quadrant == 3|quadrant == 4, as.numeric(longitude)*(-1)-lon_shift, as.numeric(longitude)+lon_shift))

# Add fleet/country
interactions <- interactions %>% 
  mutate(FLAG = gsub(" |[.]", "", FLAG)) %>% # fix extra spaces 
  left_join(fleet[,c("Fleet", "Country")], by=c("FLAG"= "Fleet")) %>%
  mutate(Country = gsub("[.]", " ", Country), 
         Country = gsub("REP", "Republic", Country), 
         Country = str_to_title(Country), 
         Country = gsub("Uk", "UK", Country), 
         Country = gsub("Eu ", "", Country), 
         Country = gsub(",China", "", Country))

# Add species data
interactions <- left_join(interactions, interactions_spp[, c("CODE", "NAME_EN")], 
                          by = c("SPECIES" = "CODE")) %>% 
  mutate(NAME_EN = str_to_sentence(NAME_EN)) %>% 
  left_join(., species[,c("Scientific_name", "English_name")], by = c("NAME_EN" = "English_name"))

# Make all species names all caps
interactions$species_sciname <- toupper(interactions$Scientific_name)
interactions$species_commonname <- toupper(interactions$NAME_EN)

# Fix species names that were not included in the ASFIS codes
interactions <- interactions %>% 
  mutate(species_commonname = ifelse(is.na(SPECIES), "OTHER SPECIES", species_commonname),
    species_sciname = ifelse(is.na(SPECIES), "OTHER SPECIES", species_sciname))

# Add gear
interactions <- left_join(interactions, gear[,c("CECode", "gear")], by=c("GEAR" = "CECode")) %>%
  mutate(gear = tolower(gear), 
         gear = gsub("  ", " ", gear), 
         gear = gsub("purse seine", "purse-seine", gear),
         gear = ifelse(GEAR == "BBOF", "baitboat (offshore)", gear))

# Add effort units
interactions <- interactions %>% 
  mutate(EFFORT_UNIT = paste0(str_to_lower(EFFORT_UNIT), "s"))

# Create/mutate relevant columns
interactions <- interactions %>% 
  mutate(rfmo = "IOTC", 
         year = YEAR, 
         month = MONTH, 
         settype = NA, 
         catch = OBSERVED_CATCH, 
         time_period = "monthly", 
         country = Country,
         effort = OBSERVED_EFFORT,
         effort_units = EFFORT_UNIT,
         catch_units = "count",
         measurement = NA, 
         measurement_unit = NA, 
         measurement_method = NA, 
         quality_code = NA, 
         was_generated = "no", 
         sources = "IOTC-2020-WPEB16-DATA12_CE")

# Select relevant columns 
interactions <- interactions[,c("rfmo", "year", "month", "time_period", "country", "gear", "settype", "longitude", "latitude", "effort", "effort_units", "species_commonname", "species_sciname", "catch", "catch_units", "measurement", "measurement_unit", "measurement_method", "spatial_notes", "quality_code", "was_generated", "sources")]

# Generate quarterly data (aggregates)
quart_agg <- interactions %>% 
  filter(time_period == "monthly") %>%
  mutate(month = ceiling(month/3),
         month = ifelse(month == 1, 1, # fix so each quarter == first month of quarter
                        ifelse(month == 2, 4, 
                               ifelse(month == 3, 7, 10))),
         time_period = "quarterly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(interactions), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Generate yearly data (aggregates)
yearly_agg <- interactions %>% 
  filter(time_period == "monthly") %>%
  mutate(month = NA,
         time_period = "yearly", 
         was_generated = "yes") %>% 
  group_by_at(setdiff(names(interactions), c("catch", "effort"))) %>%
  summarise(catch = sum(catch), 
            effort = sum(effort))

# Combine raw with generated data
interactions <- interactions %>% 
  bind_rows(quart_agg) %>% 
  bind_rows(yearly_agg)

# Save File
filename <- "interactions.csv"
write.csv(interactions, paste0(save_location, filename), row.names = FALSE)

# Clean up
remove(interactions, quart_agg, yearly_agg)
```

[Back To Top](#top)

# Combining Files
```{r echo=TRUE}
# Find list of cleaned data
files_list <- list.files(save_location, pattern="csv")

# Make sure it doesn't include the 'all' dataset
files_list <- files_list[!grepl("all", files_list)]

# Dataframe to hold combined data
combined_dat <- NULL

# Loop through files and save to master dataframe
for(file in files_list) {
  temp <- read.csv(paste0(save_location, file))
  combined_dat <- rbind(combined_dat, temp)
}

# Remove instances of NA for lat or lon (some issues with grid in sfskh)
combined_dat <- combined_dat %>% 
  filter(!is.na(latitude) & !is.na(longitude))

# Filter so only 1x1 or 5x5 data are included
combined_dat <- combined_dat %>% 
  filter(spatial_notes %in% c("center of 5x5 cell", "center of 1x1 cell") & 
           grepl("purse-seine|purse seine|longline|long line|long-line", gear))

# Save final data file for this RFMO
filename <- "iotc-all.csv"
write.csv(combined_dat, paste0(outputs_location, filename), row.names = FALSE)
```
[Back To Top](#top)